### 2020-12-21 README.Rmd
revise template to our project

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# What do incoming university students believe about open science practices in psychology?

<!-- badges: start -->
<!-- badges: end -->

This project explores incoming students' knowledge about, and attitudes toward,
open science practices. This project is led by Michael Philipp, Matthew Williams, 
and Jennifer Beaudry.

OSF page for this project: https://osf.io/gmsbn/

<!-- _Notes from Dani's template:_ -->

<!-- This is a template for a new research project. Should contains the following -->
<!-- folders: [analysis](./analysis), [data](./data), [experiments](./experiments), -->
<!-- [other](./other), [preprocessing](./preprocessing). Optionally, it may also  -->
<!-- include a folder `models` containing modelling files, and a `writeup` folder -->
<!-- containing the project writeup.  -->

<!-- If there are other resouces (e.g., OSF folders, Overleaf documents, Google docs) -->
<!-- that are linked to the project, they should be listed in this document. -->