### 2020-12-21 analysis/untitled.Rmd

---
title: "Untitled"
author: 'Brandon McMurtrie, ID: 18036371'
date: "09/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(tidyverse)
library(Rmisc)
library(rcompanion)

```

```{r loading data}
frames <- read.csv(file = "open_science.csv")
```

### Critical Counter-norm

```{r critical cnorm}

critical_cnorm.NA.RM <- na.omit(frames$critical_cnorm)

sd(critical_cnorm.NA.RM)

CI(critical_cnorm.NA.RM, ci = 0.95)

agree_critical_cnorm <- critical_cnorm.NA.RM >= 4
p.agree_critical_cnorm <- mean(agree_critical_cnorm)

disagree_critical_cnorm <- critical_cnorm.NA.RM <=2
p.disagree_critical_cnorm <- mean(disagree_critical_cnorm)

p.agree_critical_cnorm

p.disagree_critical_cnorm

length(critical_cnorm.NA.RM)

```

| Mean  | SD    | 95% CI    | P agree | P disagree |
|:-----:|:-----:|:---------:|:-------:|:---------:|
| 2.41 | 1.06 | 2.27, 2.54 | 0.24 |0.68|

### Critical Norm
```{r critical_norm}
critical_norm.NA.RM <- na.omit(frames$critical_norm)

sd(critical_norm.NA.RM)

CI(critical_norm.NA.RM, ci = 0.95)

agree_critical_norm <- critical_norm.NA.RM >= 4
p.agree_critical_norm <- mean(agree_critical_norm)

disagree_critical_norm <- critical_norm.NA.RM <=2
p.disagree_critical_norm <- mean(disagree_critical_norm)

p.agree_critical_norm
p.disagree_critical_norm

length(critical_norm.NA.RM)
```

| Mean  | SD    | 95% CI    | P agree | P disagree |
|:-----:|:-----:|:---------:|:-------:|:---------:|
| 4.32 | 0.82 | 4.21,4.42 | 0.93 | 0.05  |

### Preregistration Norm

```{r Prereg_norm}
Prereg_norm.NA.RM <- na.omit(frames$Prereg_norm)

sd(Prereg_norm.NA.RM)

agree_Prereg_norm <- Prereg_norm.NA.RM >= 4
p.agree_Prereg_norm <- mean(agree_Prereg_norm)

disagree_Prereg_norm <- Prereg_norm.NA.RM <=2
p.disagree_Prereg_norm <- mean(disagree_Prereg_norm)

CI(Prereg_norm.NA.RM, ci = 0.95)

p.agree_Prereg_norm
p.disagree_Prereg_norm

length(Prereg_norm.NA.RM)
```

| Mean  | SD    | 95% CI    | P agree | P disagree |
|:-----:|:-----:|:---------:|:-------:|:---------:|
| 3.96 | 1.07 | 3.82, 4.10 | 0.75 | 0.12 |

### Preregistration Counter-norm

```{r Prereg_cnorm}
Prereg_cnorm.NA.RM <- na.omit(frames$Prereg_cnorm)

sd(Prereg_cnorm.NA.RM)

CI(Prereg_cnorm.NA.RM, ci = 0.95)

agree_Prereg_cnorm <- Prereg_cnorm.NA.RM >= 4
p.agree_Prereg_cnorm <- mean(agree_Prereg_cnorm)

disagree_Prereg_cnorm <- Prereg_cnorm.NA.RM <=2
p.disagree_Prereg_cnorm <- mean(disagree_Prereg_cnorm)

p.agree_Prereg_cnorm
p.disagree_Prereg_cnorm

length(Prereg_cnorm.NA.RM)
```

| Mean  | SD    | 95% CI    | P agree | P disagree |
|:-----:|:-----:|:---------:|:-------:|:---------:|
| 2.74 | 1.24 | 2.58, 2.90 | 0.32 | 0.50 |

### Regreport Norm

```{r Regreport norm}
RegReport_norm.NA.RM <- na.omit(frames$RegReport_norm)

sd(RegReport_norm.NA.RM)

CI(RegReport_norm.NA.RM, ci = 0.95)

agree_RegReport_norm <- RegReport_norm.NA.RM >= 4
p.agree_RegReport_norm <- mean(agree_RegReport_norm)

disagree_RegReport_norm <- RegReport_norm.NA.RM <=2
p.disagree_RegReport_norm <- mean(disagree_RegReport_norm)

p.agree_RegReport_norm
p.disagree_RegReport_norm

length(RegReport_norm.NA.RM)
```

|Question| Mean  | SD    | 95% CI    | P agree | P disagree |
|:-----:|:-----:|:-----:|:---------:|:-------:|:---------:|
|Reg_report| 3.35 | 1.26 | 3.19, 3.52 | 0.56 | 0.27 |

### Regreport Counter-norm

```{r Regreport cnorm}
RegReport_cnorm.NA.RM <- na.omit(frames$RegReport_cnorm)

sd(RegReport_cnorm.NA.RM)

CI(RegReport_cnorm.NA.RM, ci = 0.95)

agree_RegReport_cnorm <- RegReport_cnorm.NA.RM >= 4
p.agree_RegReport_cnorm <- mean(agree_RegReport_cnorm)

disagree_RegReport_cnorm <- RegReport_cnorm.NA.RM <=2
p.disagree_RegReport_cnorm <- mean(disagree_RegReport_cnorm)

p.agree_RegReport_cnorm
p.disagree_RegReport_cnorm

length(RegReport_cnorm.NA.RM)
```

|Question| Mean  | SD    | 95% CI    | P agree | P disagree |
|:-----:|:-----:|:-----:|:---------:|:-------:|:---------:|
|Reg_report_cnorm| 3.56 | 1.20 | 3.41, 3.71 | 0.59 | 0.26 |

### p hacking counter-norm

```{r phack_cnorm}
phack_cnorm.NA.RM <- na.omit(frames$phack_cnorm)

sd(phack_cnorm.NA.RM)

CI(phack_cnorm.NA.RM, ci = 0.95)

agree_phack_cnorm <- phack_cnorm.NA.RM >= 4
p.agree_phack_cnorm <- mean(agree_phack_cnorm)

disagree_phack_cnorm <- phack_cnorm.NA.RM <=2
p.disagree_phack_cnorm <- mean(disagree_phack_cnorm)

p.agree_phack_cnorm
p.disagree_phack_cnorm

length(phack_cnorm.NA.RM)
```

|Question| Mean  | SD    | 95% CI    | P agree | P disagree |
|:-----:|:-----:|:-----:|:---------:|:-------:|:---------:|
|phack_cnorm| 2.12 | 1.27 | 1.96, 2.29 | 0.19 | 0.75 |

### p hacking Norm

```{r phack_norm}
phack_norm.NA.RM <- na.omit(frames$phack_norm)

sd(phack_norm.NA.RM)

CI(phack_norm.NA.RM, ci = 0.95)

agree_phack_norm <- phack_norm.NA.RM >= 4
p.agree_phack_norm <- mean(agree_phack_norm)

disagree_phack_norm <- phack_norm.NA.RM <=2
p.disagree_phack_norm <- mean(disagree_phack_norm)

p.agree_phack_norm
p.disagree_phack_norm

length(phack_norm.NA.RM)
```

|Question| Mean  | SD    | 95% CI    | P agree | P disagree |
|:-----:|:-----:|:-----:|:---------:|:-------:|:---------:|
|phack_norm| 4.33 | 0.90 | 4.22, 4.45 | 0.87 | 0.08 |

### Hark counter-norm

```{r HARK cnorm}
HARK_cnorm.NA.RM <- na.omit(frames$HARK_cnorm)

sd(HARK_cnorm.NA.RM)

CI(HARK_cnorm.NA.RM, ci = 0.95)

agree_HARK_cnorm <- HARK_cnorm.NA.RM >= 4
p.agree_HARK_cnorm <- mean(agree_HARK_cnorm)

disagree_HARK_cnorm <- HARK_cnorm.NA.RM <=2
p.disagree_HARK_cnorm <- mean(disagree_HARK_cnorm)

p.agree_HARK_cnorm
p.disagree_HARK_cnorm

length(HARK_cnorm.NA.RM)
```

|Question| Mean  | SD    | 95% CI    | P agree | P disagree |
|:-----:|:-----:|:-----:|:---------:|:-------:|:---------:|
|HARK_cnorm| 1.65 | 1.07 | 1.52, 1.79 | 0.11 | 0.82 |

### Hark Norm

```{r HARK Norm}
HARK_norm.NA.RM <- na.omit(frames$HARK_norm)

sd(HARK_norm.NA.RM)

CI(HARK_norm.NA.RM, ci = 0.95)

agree_HARK_norm <- HARK_norm.NA.RM >= 4
p.agree_HARK_norm <- mean(agree_HARK_norm)

disagree_HARK_norm <- HARK_norm.NA.RM <=2
p.disagree_HARK_norm <- mean(disagree_HARK_norm)

p.agree_HARK_norm
p.disagree_HARK_norm

length(HARK_norm.NA.RM)
```

|Question| Mean  | SD    | 95% CI    | P agree | P disagree |
|:-----:|:-----:|:-----:|:---------:|:-------:|:---------:|
|HARK_norm| 3.88 | 1.09 | 3.74, 4.02 | 0.70 | 0.12 |

### InfoForRep Norm

```{r InfoForRep_norm}
InfoForRep_norm.NA.RM <- na.omit(frames$InfoForRep_norm)

sd(InfoForRep_norm.NA.RM)

CI(InfoForRep_norm.NA.RM, ci = 0.95)

agree_InfoForRep_norm <- InfoForRep_norm.NA.RM >= 4
p.agree_InfoForRep_norm <- mean(agree_InfoForRep_norm)

disagree_InfoForRep_norm <- InfoForRep_norm.NA.RM <=2
p.disagree_InfoForRep_norm <- mean(disagree_InfoForRep_norm)

p.agree_InfoForRep_norm
p.disagree_InfoForRep_norm

length(InfoForRep_norm.NA.RM)
```

|Question| Mean  | SD    | 95% CI    | P agree | P disagree |
|:-----:|:-----:|:-----:|:---------:|:-------:|:---------:|
|InfoForRep_norm| 4.42 | 0.99 | 4.29, 4.55 | 0.88 | 0.08 |

### InfoForRep Counter-norm

```{r InfoForRep_cnorm}
InfoForRep_cnorm.NA.RM <- na.omit(frames$InfoForRep_cnorm)

sd(InfoForRep_cnorm.NA.RM)

CI(InfoForRep_cnorm.NA.RM, ci = 0.95)

agree_InfoForRep_cnorm <- InfoForRep_cnorm.NA.RM >= 4
p.agree_InfoForRep_cnorm <- mean(agree_InfoForRep_cnorm)

disagree_InfoForRep_cnorm <- InfoForRep_cnorm.NA.RM <=2
p.disagree_InfoForRep_cnorm <- mean(disagree_InfoForRep_cnorm)

p.agree_InfoForRep_cnorm
p.disagree_InfoForRep_cnorm

length(InfoForRep_cnorm.NA.RM)
```

|Question| Mean  | SD    | 95% CI    | P agree | P disagree |
|:-----:|:-----:|:-----:|:---------:|:-------:|:---------:|
|InfoForRep_cnorm| 1.52 | 0.95 | 1.40, 1.64 | 0.07 | 0.87 |

### Preprint Norm

```{r Preprint_norm}
Preprint_norm.NA.RM <- na.omit(frames$Preprint_norm)

sd(Preprint_norm.NA.RM)

CI(Preprint_norm.NA.RM, ci = 0.95)

agree_Preprint_norm <- Preprint_norm.NA.RM >= 4
p.agree_Preprint_norm <- mean(agree_Preprint_norm)

disagree_Preprint_norm <- Preprint_norm.NA.RM <=2
p.disagree_Preprint_norm <- mean(disagree_Preprint_norm)

p.agree_Preprint_norm
p.disagree_Preprint_norm

length(Preprint_norm.NA.RM)
```

|Question| Mean  | SD    | 95% CI    | P agree | P disagree |
|:-----:|:-----:|:-----:|:---------:|:-------:|:---------:|
|Preprint_norm| 1.61 | 0.78 | 1.51, 1.71 | 0.03 | 0.90 |

### Preprint Counter-Norm

```{r Preprint_cnorm}
Preprint_cnorm.NA.RM <- na.omit(frames$Preprint_cnorm)

sd(Preprint_cnorm.NA.RM)

CI(Preprint_cnorm.NA.RM, ci = 0.95)

agree_Preprint_cnorm <- Preprint_cnorm.NA.RM >= 4
p.agree_Preprint_cnorm <- mean(agree_Preprint_cnorm)

disagree_Preprint_cnorm <- Preprint_cnorm.NA.RM <=2
p.disagree_Preprint_cnorm <- mean(disagree_Preprint_cnorm)

p.agree_Preprint_cnorm
p.disagree_Preprint_cnorm

length(Preprint_cnorm.NA.RM)
```

|Question| Mean  | SD    | 95% CI    | P agree | P disagree |
|:-----:|:-----:|:-----:|:---------:|:-------:|:---------:|
|Preprint_cnorm| 4.36 | 0.81 | 4.26, 4.46 | 0.88 | 0.03 |

### Open Materials Norm

```{r OpenMaterials_norm}
OpenMaterials_norm.NA.RM <- na.omit(frames$OpenMaterials_norm)

sd(OpenMaterials_norm.NA.RM)

CI(OpenMaterials_norm.NA.RM, ci = 0.95)

agree_OpenMaterials_norm <- OpenMaterials_norm.NA.RM >= 4
p.agree_OpenMaterials_norm <- mean(agree_OpenMaterials_norm)

disagree_OpenMaterials_norm <- OpenMaterials_norm.NA.RM <=2
p.disagree_OpenMaterials_norm <- mean(disagree_OpenMaterials_norm)

p.agree_OpenMaterials_norm
p.disagree_OpenMaterials_norm

length(OpenMaterials_norm.NA.RM)
```

|Question| Mean  | SD    | 95% CI    | P agree | P disagree |
|:-----:|:-----:|:-----:|:---------:|:-------:|:---------:|
|OpenMaterials_norm| 3.68 | 1.03 | 3.55, 3.82 | 0.66 | 0.15 |

### Open Materials Counter-norm

```{r OpenMaterials_cnorm}
OpenMaterials_cnorm.NA.RM <- na.omit(frames$OpenMaterials_cnorm)

sd(OpenMaterials_cnorm.NA.RM)

CI(OpenMaterials_cnorm.NA.RM, ci = 0.95)

agree_OpenMaterials_cnorm <- OpenMaterials_cnorm.NA.RM >= 4
p.agree_OpenMaterials_cnorm <- mean(agree_OpenMaterials_cnorm)

disagree_OpenMaterials_cnorm <- OpenMaterials_cnorm.NA.RM <=2
p.disagree_OpenMaterials_cnorm <- mean(disagree_OpenMaterials_cnorm)

p.agree_OpenMaterials_cnorm
p.disagree_OpenMaterials_cnorm

length(OpenMaterials_cnorm.NA.RM)
```

|Question| Mean  | SD    | 95% CI    | P agree | P disagree |
|:-----:|:-----:|:-----:|:---------:|:-------:|:---------:|
|OpenMaterials_cnorm| 2.5 | 1.15 | 2.35, 2.65 | 0.23 | 0.60 |

### Open Data Norm

```{r OpenData_norm}
OpenData_norm.NA.RM <- na.omit(frames$OpenData_norm)

sd(OpenData_norm.NA.RM)

CI(OpenData_norm.NA.RM, ci = 0.95)

agree_OpenData_norm <- OpenData_norm.NA.RM >= 4
p.agree_OpenData_norm <- mean(agree_OpenData_norm)

disagree_OpenData_norm <- OpenData_norm.NA.RM <=2
p.disagree_OpenData_norm <- mean(disagree_OpenData_norm)

p.agree_OpenData_norm
p.disagree_OpenData_norm

length(OpenData_norm.NA.RM)
```

|Question| Mean  | SD    | 95% CI    | P agree | P disagree |
|:-----:|:-----:|:-----:|:---------:|:-------:|:---------:|
|OpenData_norm| 2.65 | 1.21 | 2.49, 2.80 | 0.26 | 0.49 |

### Open Data Counter-norm

```{r OpenData_cnorm}
OpenData_cnorm.NA.RM <- na.omit(frames$OpenData_cnorm)

sd(OpenData_cnorm.NA.RM)

CI(OpenData_cnorm.NA.RM, ci = 0.95)

agree_OpenData_cnorm <- OpenData_cnorm.NA.RM >= 4
p.agree_OpenData_cnorm <- mean(agree_OpenData_cnorm)

disagree_OpenData_cnorm <- OpenData_cnorm.NA.RM <=2
p.disagree_OpenData_cnorm <- mean(disagree_OpenData_cnorm)

p.agree_OpenData_cnorm
p.disagree_OpenData_cnorm

length(OpenData_cnorm.NA.RM)
```

|Question| Mean  | SD    | 95% CI    | P agree | P disagree |
|:-----:|:-----:|:-----:|:---------:|:-------:|:---------:|
|OpenData_cnorm| 3.49 | 1.25 | 3.33, 3.65 | 0.55 | 0.28 |

### Open Acc Norm

```{r OpenAcc_norm}
OpenAcc_norm.NA.RM <- na.omit(frames$OpenAcc_norm)

sd(OpenAcc_norm.NA.RM)

CI(OpenAcc_norm.NA.RM, ci = 0.95)

agree_OpenAcc_norm <- OpenAcc_norm.NA.RM >= 4
p.agree_OpenAcc_norm <- mean(agree_OpenAcc_norm)

disagree_OpenAcc_norm <- OpenAcc_norm.NA.RM <=2
p.disagree_OpenAcc_norm <- mean(disagree_OpenAcc_norm)

p.agree_OpenAcc_norm
p.disagree_OpenAcc_norm

length(OpenAcc_norm.NA.RM)
```

|Question| Mean  | SD    | 95% CI    | P agree | P disagree |
|:-----:|:-----:|:-----:|:---------:|:-------:|:---------:|
|OpenAcc_norm| 3.85 | 0.90 | 3.74, 3.97 | 0.70 | 0.07 |

### Open Acc Counter-norm

```{r OpenAcc_cnorm}
OpenAcc_cnorm.NA.RM <- na.omit(frames$OpenAcc_cnorm)

sd(OpenAcc_cnorm.NA.RM)

CI(OpenAcc_cnorm.NA.RM, ci = 0.95)

agree_OpenAcc_cnorm <- OpenAcc_cnorm.NA.RM >= 4
p.agree_OpenAcc_cnorm <- mean(agree_OpenAcc_cnorm)

disagree_OpenAcc_cnorm <- OpenAcc_cnorm.NA.RM <=2
p.disagree_OpenAcc_cnorm <- mean(disagree_OpenAcc_cnorm)

p.agree_OpenAcc_cnorm
p.disagree_OpenAcc_cnorm

length(OpenAcc_cnorm.NA.RM)
```

|Question| Mean  | SD    | 95% CI    | P agree | P disagree |
|:-----:|:-----:|:-----:|:---------:|:-------:|:---------:|
|OpenAcc_cnorm| 2.53 | 0.92 | 2.42, 2.65 | 0.13 | 0.48 |

### Norms in Practice 1

``` {r practices 1}

mean(frames$Practices_1, na.rm = TRUE)

sd(frames$Practices_1, na.rm = TRUE)

t.test(frames$Practices_1)

```

### Norms in Practice 2

``` {r practices 2}

mean(frames$Practices_2, na.rm = TRUE)

sd(frames$Practices_2, na.rm = TRUE)

t.test(frames$Practices_2)

```

### Norms in Practice 3

``` {r practices 3}

mean(frames$Practices_3, na.rm = TRUE)

sd(frames$Practices_3, na.rm = TRUE)

t.test(frames$Practices_3)

```

### Norms in Practice 4

``` {r practices 4}

mean(frames$Practices_4, na.rm = TRUE)

sd(frames$Practices_4, na.rm = TRUE)

t.test(frames$Practices_4)

```

### Norms in Practice 5

``` {r practices 5}

mean(frames$Practices_5, na.rm = TRUE)

sd(frames$Practices_5, na.rm = TRUE)

t.test(frames$Practices_5)

```

### Norms in Practice 6

``` {r practices 6}

mean(frames$Practices_6, na.rm = TRUE)

sd(frames$Practices_6, na.rm = TRUE)

t.test(frames$Practices_6)

```

### Norms in Practice 7

``` {r practices 7}

mean(frames$Practices_7, na.rm = TRUE)

sd(frames$Practices_7, na.rm = TRUE)

t.test(frames$Practices_7)

```

### Norms in Practice 8

``` {r practices 8}

mean(frames$Practices_8, na.rm = TRUE)

sd(frames$Practices_8, na.rm = TRUE)

t.test(frames$Practices_8)

```

### Norms in Practice 9

``` {r practices 9}

mean(frames$Practices_9, na.rm = TRUE)

sd(frames$Practices_9, na.rm = TRUE)

t.test(frames$Practices_9)

```

### Norms in Practice 10

``` {r practices 10}

mean(frames$Practices_10, na.rm = TRUE)

sd(frames$Practices_10, na.rm = TRUE)

t.test(frames$Practices_10)

```

### Bivariate Analyses from Preregistration.

```{r correlation between critcal norm and counter norm}
cor.test(frames$critical_norm, frames$critical_cnorm)

cor(frames$critical_norm, frames$critical_cnorm, use = "complete.obs", method = "pearson")

```

```{r correlation between Preregistration norm and counter norm}
cor.test(frames$Prereg_norm, frames$Prereg_cnorm)

cor(frames$Prereg_norm, frames$Prereg_cnorm, use = "complete.obs", method = "pearson")

```

```{r correlation between Regreport norm and counter norm}
cor.test(frames$RegReport_norm, frames$RegReport_cnorm)

cor(frames$RegReport_norm, frames$RegReport_cnorm, use = "complete.obs", method = "pearson")

```

```{r correlation between P hacking norm and counter norm}

cor.test(frames$phack_norm, frames$phack_cnorm)

cor(frames$phack_norm, frames$phack_cnorm, use = "complete.obs", method = "pearson")

```

```{r correlation between HARK norm and counter norm}
cor.test(frames$HARK_norm, frames$HARK_cnorm)

cor(frames$HARK_norm, frames$HARK_cnorm, use = "complete.obs", method = "pearson")

```

```{r correlation between InfoForRep norm and counter norm}
cor.test(frames$InfoForRep_norm, frames$InfoForRep_cnorm)

cor(frames$InfoForRep_norm, frames$InfoForRep_cnorm, use = "complete.obs", method = "pearson")

```

```{r correlation between PrePrint norm and counter norm}
cor.test(frames$Preprint_norm, frames$Preprint_cnorm)

cor(frames$Preprint_norm, frames$Preprint_cnorm, use = "complete.obs", method = "pearson")

```

```{r correlation between Open Materials norm and counter norm}
cor.test(frames$OpenMaterials_norm, frames$OpenMaterials_cnorm)

cor(frames$OpenMaterials_norm, frames$OpenMaterials_cnorm, use = "complete.obs", method = "pearson")

```

```{r correlation between Open Data norm and counter norm}
cor.test(frames$OpenData_norm, frames$OpenData_cnorm)

cor(frames$OpenData_norm, frames$OpenData_cnorm, use = "complete.obs", method = "pearson")

```

```{r correlation between Open Acc norm and counter norm}
cor.test(frames$OpenAcc_norm, frames$OpenAcc_cnorm)

cor(frames$OpenAcc_norm, frames$OpenAcc_cnorm, use = "complete.obs", method = "pearson")

```

### Difference in means between responses to norm and counter-norm, and 95% CI

```{r t test critical}

t.test(frames$critical_norm, frames$critical_cnorm, paired = TRUE, alternative = "two.sided")

```

```{r t test prereg norm}

t.test(frames$Prereg_norm, frames$Prereg_cnorm, paired = TRUE, alternative = "two.sided")

```

```{r t test Reg Report}

t.test(frames$RegReport_norm, frames$RegReport_cnorm, paired = TRUE, alternative = "two.sided")

```

```{r t test p hack}

t.test(frames$phack_norm, frames$phack_cnorm, paired = TRUE, alternative = "two.sided")

```

```{r t test HARKING}

t.test(frames$HARK_norm, frames$HARK_cnorm, paired = TRUE, alternative = "two.sided")

```

```{r t test InfoForRep}

t.test(frames$InfoForRep_norm, frames$InfoForRep_cnorm, paired = TRUE, alternative = "two.sided")

```

```{r t test Preprint}

t.test(frames$Preprint_norm, frames$Preprint_cnorm, paired = TRUE, alternative = "two.sided")

```

```{r t test OpenMaterials}

t.test(frames$OpenMaterials_norm, frames$OpenMaterials_cnorm, paired = TRUE, alternative = "two.sided")

```

```{r t test Open Data}

t.test(frames$OpenData_norm, frames$OpenData_cnorm, paired = TRUE, alternative = "two.sided")

```

```{r t test OpenAcc}

t.test(frames$OpenAcc_norm, frames$OpenAcc_cnorm, paired = TRUE, alternative = "two.sided")

```