### 2021-03-09 docs/index.Rmd
Tightening up third paragraph
Rewriting a few parts of this paragraph to help the flow a bit.

---
title: "Project Documentation"
author: "Danielle Navarro"
date: "23/08/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The "docs" folder on GitHub is special: if your project is set up to allow it, any files in this folder will appear as a GitHub Pages website. For example, this template is located at https://github.com/djnavarro/newproject, and because I have enabled GitHub pages for this project all the files in this folder are hosted at https://djnavarro.github.io/newproject. It is very often useful to have a public facing website associated with a project that gives people an overview of the whole thing, and this is an easy way to do it.

To help you get started, GitHub have a [nice walkthrough](https://pages.github.com/#project-site) showing you how to enable GitHub Pages for a specific project. There is also a more detailed tutorial [here](https://help.github.com/en/articles/configuring-a-publishing-source-for-github-pages)